---
title: "Polyglot Analysis: Stata + R"
author: "Your Name"
date: today
format: html
engine: knitr
---

## Overview

This document demonstrates a **Stata + R polyglot** Quarto workflow. Stata handles
data loading and statistical modelling; R handles visualisation via ggplot2. Data
is passed between the two languages through a shared DTA file.

> **Prerequisites**
> - R ≥ 4.0 with packages: `knitr`, `haven`, `ggplot2`
>   (`install.packages(c("knitr", "haven", "ggplot2"))`)
> - Stata 17 or later
> - Update `engine.path` below to match your Stata installation

---

## Setup

```{r setup, include=FALSE}
library(haven)
library(ggplot2)

knitr::opts_chunk$set(
  echo    = TRUE,
  comment = "",
  engine.path = list(
    stata = "/path/to/stata"
    # Linux / macOS : "/usr/local/stata18/stata-mp"
    # Windows       : "C:/Program Files/Stata18/StataMP-64.exe"
  )
)
```

---

## Step 1: Load and clean data (Stata)

```{stata load}
sysuse auto, clear
label variable price  "Price (USD)"
label variable mpg    "Mileage (mpg)"
label variable weight "Weight (lbs)"
label variable foreign "Origin (0=Domestic, 1=Foreign)"
describe
```

## Step 2: Regression model (Stata)

```{stata regression}
regress price mpg weight i.foreign
estimates store ols
```

## Step 3: Export for R (Stata)

```{stata export}
save auto_for_r.dta, replace
```

## Step 4: Visualise in R

R reads the DTA file written by Stata and produces a ggplot2 chart.

```{r visualise, fig.cap="Price versus mileage, coloured by car origin"}
df <- read_dta("auto_for_r.dta")

# Decode the `foreign` value label into a readable factor
df$origin <- factor(df$foreign,
                    levels = c(0, 1),
                    labels = c("Domestic", "Foreign"))

ggplot(df, aes(x = mpg, y = price, colour = origin)) +
  geom_point(size = 2.5, alpha = 0.8) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 0.8) +
  scale_colour_manual(values = c("Domestic" = "#1f77b4",
                                  "Foreign"  = "#d62728")) +
  labs(
    title  = "Price vs. Mileage by Car Origin",
    x      = "Mileage (mpg)",
    y      = "Price (USD)",
    colour = "Origin"
  ) +
  theme_minimal()
```

## Step 5: Summary table in R

```{r summary-table}
agg <- aggregate(cbind(price, mpg, weight) ~ origin, data = df, FUN = mean)
knitr::kable(agg, digits = 1, caption = "Mean values by car origin")
```

---

## Cleanup

```{r cleanup, include=FALSE}
file.remove("auto_for_r.dta")
```

---

## Positron IDE Note

When opening this file in Positron with the `positron-stata` extension active,
Stata cells execute in the running Stata session and R cells execute in the
running R session. No `engine.path` configuration is needed for the interactive
preview path.
